---
title: "Analysis of COVID-era Apple mobility Data"
author: "Alaa Alazzeh"
date: "2021-11-28"
output: html_document
bibliography: references.bib
params: 
  state: "New York"
  data: "data/raw_data/applemobilitytrends-2021-09-29.csv"
  seqdata: "output/sars_parsing_data.txt"
---

```{r setup, include=FALSE}
# when knitting keeping the echo and message to TRUE allows you to see the
# code, messages, and troubleshooting messages that helps you while developing
# the code that you. If you changed the echo and message to FALSE then it would
# produce a clean and well formatted document without the extra notes
knitr::opts_chunk$set(echo = TRUE, message = TRUE)
```

## Setup

```{r set-parameters}
# set the state to work with
state_to_analyze <- params$state

# sanitizing the state name for output files
state_no_spaces <- gsub(state_to_analyze,
                        pattern = " ",
                        replacement = "_")

#set the name of the global data file to use
global_file_to_process <- params$data

# extract core name for output file naming use
core_name <- tools::file_path_sans_ext(basename(global_file_to_process))
```

```{r load-packages-functions}
# load the packages needed
library("dplyr")
library("readr")
library("ggplot2")
library("tidyr")
library("usethis")
library("lubridate")
library("devtools")

# load functions needed
source("code/functions/subset_data_to_a_state.R")
source("code/functions/count_up_counties_cities.R")
source("code/functions/wide_subset_data_to_long.R")
```

## Background and Objectives

This markdown file will use the `ggplot2` [@ggplot2] and `dplyr` [@dplyr] packages to analyze the Apple maps mobility data from the time before the COVID-19 pandemic through the initial shutdowns and the continuing changes over the course of the year.

The goal will be to create a subset to a particular state, tally up the number of the cities and counties in that state with mobility data, and then to make several plots to visualize those data subsets.

The state that this document will analyze is `r state_to_analyze`.

## Data Processing 

```{r susbet-state}
# subset out to the chosen state using the variables set above
state_subset <- subset_data_to_a_state(
  input_file_name = global_file_to_process,
  state_to_subset = state_to_analyze)
```

```{r make-subset-long}
# convert the subsetted state data from wide to long format
long_subset <- wide_subset_data_to_long(
  input_file_name = paste0("output/01_subset_state_data/",
                        core_name,
                        "_",
                        state_no_spaces,
                        ".csv"))
```

```{r tallied-subset}
# tally up the number of cities and counties in the state that have mobility
# data available
tallied_subset <- count_up_counties_cities(
  input_file_name = paste0("output/01_subset_state_data/",
                        core_name,
                        "_",
                        state_no_spaces,
                        ".csv"))
```

## Figures

```{r plot-tallied}
# create a graphed barplot of the tallied data to include in knitted
#document, and save that figure to the output/figures directory as a png
tallied_plot <- ggplot(data = tallied_subset,
                       aes(x = geo_type,
                           y = n,
                           fill = transportation_type)) +
  geom_col(position = position_dodge()) +
  labs(title = paste("Number of cities and/or counties in",
                     state_to_analyze,
                     "with mobility data"),
       x = "Type of area",
       y = "Number of areas")

ggsave(plot = tallied_plot,
       filename = paste0("output/figures/",
                         core_name,
                         "_",
                         state_no_spaces,
                         "_talied_plot.png"))

# R requires you to write out the name of the plot so it can
# produce it at the end of the code
tallied_plot
```

```{r plotting-time-series}
# produce a line plot of relative driving mobility across the state with date
# on the x axis and mean relative mobility on the y axis

# This allows for only driving data to be selected. From selected data
#  we average all of the relative mobilities from a specific date to
# give what is now called mean_mobility. Then we create a plot right
# after. Note: If we go into ggplot after dplyr then we dont need the
# first data argument because it is handed it from the end of the dplyr # chain
timeseries_plot <- long_subset %>%
  filter(transportation_type == "driving") %>%
  group_by(date) %>%
  summarize(mean_mobility = mean(relative_mobility)) %>%
  ggplot(aes(x = lubridate::ymd(date),
             y = mean_mobility)) +
geom_line() +
  labs(title = paste("Statewide mean realtive mobility driving levels in",
                     state_to_analyze,
                     "during COVID"),
       x = "Date",
       y = "Mean relatibe mobility")

ggsave(plot = timeseries_plot,
       filename = paste0("output/figures/",
                         core_name,
                         "_",
                         state_no_spaces,
                         "_timeseries_plot.png"))

# R requires you to write out the name of the plot so it can
# produce it at the end of the code
timeseries_plot
```

```{r scatter-plot}
# Produce a scatterplot of all of the relative mobility data for driving
# for any given state
scatter_plot <- long_subset %>%
  filter(transportation_type == "driving") %>%
  group_by(date) %>%
  summarize(mean_mobility = mean(relative_mobility)) %>%
  ggplot(aes(x = lubridate::ymd(date),
                           y = mean_mobility)) +
  geom_point() +
  labs(title = paste("Scatterplot of mean relative driving mobility data for",
                     state_to_analyze,
                     "during COVID"),
       x = "Date",
       y = "Mean relatibe mobility")

ggsave(plot = scatter_plot,
       filename = paste0("output/figures/",
                         core_name,
                         "_",
                         state_no_spaces,
                         "_scatter_plot.png"))

scatter_plot
```

```{r plot-sequence-data, fig.height-12}
# read in the count of the sequences by country (2 columns)
seqs_by_country <- readr::read_table(params$seqdata,
                                     col_names = c("count",
                                                   "country"))

# check that we have the right number of columns
stopifnot(ncol(seqs_by_country) == 2)

# make a sideways barplot, log scale x axis, with countries on the y
# and the counts of sequences in each on the x, sorted by the count
seq_count_plot <- ggplot(data = seqs_by_country,
                         aes(x = count,
                             y = reorder(country, count))) +
        geom_col() +
        scale_x_log10() +
        labs(title = "Count of SARS-CoV-2 sequences per country",
             x = "Count of Sequences",
             y = "Country Name")

ggsave(plot = seq_count_plot,
       filename = "output/figures/seq_count_plot.png")

seq_count_plot
````

## Session Info

```{r session-info}
devtools::session_info()
```



## Sources Cited


