---
title: "Analysis of COVID-era Apple mobility Data"
author: "Alaa Alazzeh"
date: "2021-10-10"
output: html_document
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = TRUE)
```

## Setup

```{r set-parameters}
# set the state to work with
state_to_analyze <- "California"

# santizing the state name for output files
state_no_spaces <- gsub(state_to_analyze,
                        pattern = " ",
                        replacement = "_")

#set the name of the global data file to use
global_file_to_process <- "data/raw_data/applemobilitytrends-2021-09-29.csv"

# extract core name for output file naming use 
core_name <- tools::file_path_sans_ext(basename(global_file_to_process))
```

```{r load-packages-functions}
# load the packages needed 
library("dplyr")
library("readr")
library("ggplot2")
library("tidyr")
library("usethis")
library("lubridate")
library("devtools")

# load functions needed 
source("code/functions/subset_data_to_a_state.R")
source("code/functions/count_up_counties_cities.R")
source("code/functions/wide_subset_data_to_long.R")
```

## Background and Objectives

This markdown file will use the `ggplot2` [@ggplot2] and `dplyr` [@dplyr] packages to analyze the Apple maps mobility data from the time before the COVID-19 pandemic through the initial shutdowns and the continuing changes over the course of the year.

The goal will be to create a subset to a particular state, tally up the number of the cities and counties in that state with mobility data, and then to make several plots to visualize those data subsets.

The state that this document will analyze is `r state_to_analyze`


## Sources Cited


